How functional traits influence plant growth and shade tolerance across the life-cycle

### 2018-05-01 ms/ms.tex
Tweaks for technical check from journal

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

@@

\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}. Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover. During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}. As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}. In this article we outline the mechanisms by which the growth of individual plants is influenced by various functional traits, as well as plant size and the light environment.

@@

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle. The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}. Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. Based on the same decomposition of growth rates as is implemented in the {\plant} model and used below, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size. Here we extend the results of \citep{Gibert-2016} to provide a full, functioning model and use this to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size. We consider five prominent traits representing specific physiological trade-offs: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).

@@

The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in Tables \ref{tab:params_core}--\ref{tab:params_hyper}.

@@

\subsection{Classic model for mass-based relative growth rate}

@@

\subsection{Decomposition of growth rates into components}

@@

\subsection{Shade tolerance}

@@

\subsection{A functional-balance model for plant construction}

@@

\subsection{Trait-based trade-offs}

@@

\subsubsection{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.

@@

\subsubsection{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.

@@

\subsection{Model assumptions}

@@

\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
\centering
\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{growth_light_height.pdf}
\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated. The white regions indicate trait ranges that are typically observed in real systems. Figs. \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass. Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
\label{fig:growth_light_height}}
\end{SCfigure*}

\subsection{Changes in growth rate with size}

@@

\subsection{Changes in height growth rate with traits}

@@

\subsubsection{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease. The only effect of {\seed} on growth comes from changing the plant's initial size. The plots in Fig. \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size. Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area. At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases. As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle. As a result, the correlations between {\seed} and growth rate disappear among larger plants.

@@

\subsubsection{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig. \ref{fig:growth_light_height}). As $E$ increases, the optimal {\nitrogen} also increases. The invariance of the growth-trait relationship with respect to size arises as follows. The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate. Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
u}  - \frac{ \partial r_{\rm l}}{\partial
u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.

Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig. \ref{fig:growth_light_height}). As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height. Decreasing {\lma} has two impacts on height growth rate. First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$. Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover. Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects. When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate. When plants are large, the influence of {\lma} on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger. Thus, at larger sizes low {\lma} no longer increases growth rate (Fig. \ref{fig:growth_light_height}).

@@

\subsection{Changes in other growth rates with trait}

The results reported above and in Fig. \ref{fig:growth_light_height} focus on height growth rate. Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in Figs. \ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}. For each, plants were grown to a standard diameter, area, or mass. Changes in absolute and relative growth rates with traits therefore show similar patterns.

@@

\subsection{Responsiveness of growth rate to light}

The predictions in Figs. \ref{fig:growth_light_height} and \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} illustrate how traits impact on growth rate under different light environments and at different sizes. An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment. This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth. Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}). Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive. However, unlike for {\wood}, the effect appears only for the smallest size classes.

\subsection{Shade tolerance}
Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
\begin{equation}\label{eq:wplcp_2}
\bar{p}(E^*) =
      \phi c_{\rm l} +
      \left(\theta \, \rho \, \eta_c \, H\right)
        \left(b c_{\rm b}
            + c_{\rm s}\right) +
      \alpha_{\rm r1} c_{\rm r},
\end{equation}
where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}. With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$. Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$. The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively. Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig. \ref{fig:wplcp}). At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover. {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}. In addition, {\wplcp} varies with {\nitrogen}. At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e. high {\nitrogen} makes seedlings shade intolerant. At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}. All these results  match empirical patterns (Table \ref{tab:phenomena}).

@@

Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle. Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits. During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}. However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}. Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size. Moreover, we show that different traits and trade-offs generate different types of response. Combined with the available empirical evidence, these results demand a fundamental shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}. Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e. lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}. The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}. This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}. Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects. Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.

\subsection{LMA doesn't predict growth rates of large plants}

@@

\subsection{Generalising to other traits and trade-offs}

@@

\subsection{Implications for trait-based approaches}
There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.

@@

\subsection{Plasticity of traits through ontogeny}

@@

\subsection{Comparison with other frameworks}

@@

A feature distinguishing our model from most of those mentioned above is the explicit linking to trait-based trade-offs. While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants. Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model. Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig. \ref{fig:growth_light_height}). Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.

Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation. Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus. This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e. growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant. The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction. While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}. Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by ][]{Wenk-2015}. Currently, reproductive allocation is given little attention in models of ecosystem flux \citep[e.g.][]{Sitch-2008, DeKauwe-2014}. For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size. Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a plant's maximum size and growth rate in adult plants \citep[e.g.][]{Wright-2010}.

@@

\subsection{Closing remarks}

@@

\matmethods{
\vspace{-1em}

\subsection{Parameters}
Parameters were sourced mainly from \citep{Falster-2016} (see Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values). The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig. \ref{fig:conceptual}).

\subsection{Assumptions}
The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species. We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships. We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.

>>>>

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b,1]{Daniel S. Falster}
\author[c]{Remko A. Duursma}
\author[d,b]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Previous address: Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
\affil[c]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[d]{Imperial College, London, United Kingdom}

@@

\dropcap{F}unctional traits capture core differences in the strategies plants use to acquire and invest resources \citep{Westoby-2002, Wright-2004, Chave-2009}. Although most woody plants have the same basic physiological function and key resource requirements (carbon, nutrients, and water), species differ considerably in the rates at which resources are acquired, invested into different tissues, and lost via turnover. During the last two decades, trade-offs related to some prominent traits have been quantified, with values for some leaf, wood, or seed traits now available for up to 10\% of the world's 250000 plant species \citep{Cornwell-2014}. As data has accumulated, researchers are increasingly looking to traits to predict patterns in plant growth, demography, life history and performance \citep{Poorter-2008, Wright-2010, VanKleunen-2010, Adler-2014}. In this article we outline the mechanisms by which the growth of individual plants may be influenced by various functional traits, as well as plant size and the light environment.

@@

Here we show how a new mechanistic growth model -- called {\plant} \citep{Falster-2016} -- can explain diverse empirical phenomena, including a size-dependent effect of traits on growth and an effect of traits on shade tolerance (Table \ref{tab:phenomena}), and thereby offer new insights into the way traits influence key elements of plant demography across the life-cycle. The {\plant} model builds on past approaches to modelling production and allocation of biomass \citep[e.g.][]{Givnish-1988, Yokozawa-1995,Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, Gibert-2016}. Our primary focus in this article is to explain a pattern that has been gradually emerging -- that the effect of traits on plant growth is modified by plant size \citep{Ruger-2012, Iida-2014, Gibert-2016}. Based on the same decomposition of growth rates as is implemented in the {\plant} model \citep[from][]{Falster-2011, Falster-2017}, \citep{Gibert-2016} argued conceptually why the effect of traits on growth should change with size. Here we extend the results of \citep{Gibert-2016} to show, from the point of view of mass production and allocation, how and why the effect of some traits on growth rates changes with size. We consider five prominent traits, each capturing a specific physiological trade-off: seed mass ({\seed}), height at maturation ({\hmat}), leaf nitrogen content per unit leaf area ({\nitrogen}),  {\lma}, and {\wood} (Table \ref{tab:traits}).

@@

The {\plant} model builds on the widespread approach used in many vegetation models of explicitly modelling the amounts of biomass in different tissues within a plant \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001, Sitch-2008, Falster-2011, King-2011, DeKauwe-2014} (Fig. \ref{fig:conceptual}a). We consider the masses $M_i$, areas $A_i$, and diameters $D_i$ of tissues, where the subscripts indicates tissue type: l=leaf, b=bark and phloem, s=sapwood, h=heartwood, r=root, a=alive, t=total, and st=stem total (s+b+h). The total mass of living tissue is then $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}$, and the standing mass of the plant is $M_{\rm t}=M_{\rm a}+M_{\rm h}$. A summary of all variables, units, and definitions is given in Table \ref{tab:definitions}, with further details on the parameter values applied given in SI Appendix, Tables \ref{tab:params_core}--\ref{tab:params_hyper}.

@@

\subsection*{Classic model for mass-based relative growth rate}

@@

\subsection*{Decomposition of growth rates into components}

@@

\subsection*{Shade tolerance}

@@

\subsection*{A functional-balance model for plant construction}

@@

\subsection*{Trait-based trade-offs}

@@

\subsubsection*{Ontogenetic traits}  The first two traits considered modify behaviour primarily at the start and end of an individual's growth trajectory and are therefore termed ``ontogenetic traits''.

@@

\subsubsection*{Developmental traits} The remaining three traits are termed ``developmental traits'' because they moderate effects the rate of movement along an individual's growth trajectory.

@@

\subsection*{Model assumptions}

@@

\begin{SCfigure*}[\sidecaptionrelwidth][!ht]
\centering
\includegraphics[width=12.4cm,height=12.4cm, keepaspectratio]{growth_light_height.pdf}
\caption{\textbf{Effect of four traits on height growth rate for different-sized plants.}
Growth rates were simulated using the {\plant} model, applying the trade-offs describing in Table \ref{tab:traits}. Each panel shows how growth is influenced by a different trait for plants of a given height, and across a series of canopy openness values from completely open (light blue,  $E=1$) to heavily shaded (dark line, $E=0.25$). For any given value of trait and $E$, plants were grown to the desired height and their growth rate estimated. The white regions indicate trait ranges that are typically observed in real systems. In the SI Appendix, Figs. \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} show similar plots but with growth measured as stem diameter, stem area, or plant mass. Changes in trait-growth relationships are summarised in Table \ref{tab:responses}.
\label{fig:growth_light_height}}
\end{SCfigure*}

\subsection*{Changes in growth rate with size}

@@

\subsection*{Changes in height growth rate with traits}

@@

\subsubsection*{Ontogenetic traits}  Increasing {\seed} causes seedlings to be larger and the rate of seed production to decrease. The only effect of {\seed} on growth comes from changing the plant's initial size. The plots in Fig. \ref{fig:conceptual}, which show changes in growth rate with plant size, also express the expected changes in the growth of seedlings due to changes in seed size. Under similar light conditions, larger seedlings are predicted to have faster absolute growth rates (in all metrics) because of their greater total leaf area. At the same time, relative growth rate is predicted to decrease with size, as the ratio of leaf area to support mass decreases. As plants grow, differences in initial mass decrease in importance relative to other factors influencing growth through the life-cycle. As a result, the correlations between {\seed} and growth rate disappear among larger plants.

@@

\subsubsection*{Developmental traits}  The response of growth rate to changes in {\nitrogen} is relatively straightforward: there is an optimum value of {\nitrogen} that maximises height growth rate in a given light environment $E$ and does not vary with height (Fig. \ref{fig:growth_light_height}). As $E$ increases, the optimal {\nitrogen} also increases. The invariance of the growth-trait relationship with respect to size arises as follows. The direct physiological effect of {\nitrogen} is to increase the maximum potential photosynthetic rate of leaves, with a cost of higher respiration rate. Both the cost and benefits of {\nitrogen} appear within the term $\frac{ {\rm d}B}{{\rm d}t}$, implying the direction of correlation between trait and growth rate depends crucially on the change in biomass growth per {\nitrogen}.
From eqn \ref{eq:dbdt}, $\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial
u} = A_{\rm l} \, (\frac{ \partial \bar{p}(E)}{\partial
u}  - \frac{ \partial r_{\rm l}}{\partial
u} )$. As both $\bar{p}(E)$ and $r_{\rm l}$ are expressed per unit area and independent of height, the optimal value is also independent of height.

Unlike {\nitrogen}, the response of growth rate to changes in {\lma} varies strongly with plant height, with the relationship moving like a wave across the trait spectrum (Fig. \ref{fig:growth_light_height}). As a result, the value of {\lma} that optimises plant growth increases with height, while the direction of correlation between height growth rate and {\lma} shifts from negative to positive, as plants increase in height. Decreasing {\lma} has two impacts on height growth rate. First, lower {\lma} increases the leaf deployment per mass invested $(\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}})$. Second, lower {\lma} decreases biomass growth $(\frac{{\rm d}B}{{\rm d}t})$, due to increased leaf turnover. Whether lower {\lma} increases height growth depends on the relative magnitude of these two effects. When plants are small, the effect on leaf deployment rate is larger and so decreasing {\lma} increases growth rate. When plants are large, the effect on leaf deployment is diminished because the cost of building other supportive tissues (other terms in eqn \ref{eq:daldmt}) is larger. Thus, at larger sizes low {\lma} no longer increases growth rate (Fig. \ref{fig:growth_light_height}).

@@

\subsection*{Changes in other growth rates with trait}

The results reported above and in Fig. \ref{fig:growth_light_height} focus on height growth rate. Corresponding results for absolute growth rates in stem diameter (eqn \ref{eq:dDdt}), stem basal area (eqn \ref{eq:dast}), and above-ground mass (eqn \ref{eq:dMtdt}) are provided in the SI Appendix, Figs. \ref{fig:growth_light_dia}--\ref{fig:growth_light_mass}. For each, plants were grown to a standard diameter, area, or mass. Changes in absolute and relative growth rates with traits therefore show similar patterns.

@@

\subsection*{Responsiveness of growth rate to light}

The predictions in Fig. \ref{fig:growth_light_height} and Figs. \ref{fig:growth_light_dia}-\ref{fig:growth_light_mass} in the SI Appendix illustrate how traits impact on growth rate under different light environments and at different sizes. An additional outcome that arises directly from these analyses is that traits moderate the responsiveness of growth to changes in light environment. This response arises because individuals with higher potential growth rate naturally have greater potential plasticity in growth. Our results therefore support findings that species with low {\wood} increase growth more substantially with increases in light (Table \ref{tab:phenomena}). Variation in {\lma} also moderates the response of growth to changes in light, with species having the lowest {\lma} being most responsive. However, unlike for {\wood}, the effect appears only for the smallest size classes.

\subsection*{Shade tolerance}
Combining eqn \ref{eq:wplcp} with the function-balance model in Table \ref{tab:allometry} leads to the a more specific expression for calculating {\wplcp}, as the value of $E^*$ that gives
\begin{equation}\label{eq:wplcp_2}
\bar{p}(E^*) =
      \phi c_{\rm l} +
      \left(\theta \, \rho \, \eta_c \, H\right)
        \left(b c_{\rm b}
            + c_{\rm s}\right) +
      \alpha_{\rm r1} c_{\rm r},
\end{equation}
where $c_{\rm i} = \left(\frac{k_{\rm i}}{y} + r_{\rm i}\right)$ for $i=l,s,b,r$.
Eqn \ref{eq:wplcp_2} indicates {\wplcp} will increase approximately linearly with $H$ and potentially vary with {\nitrogen}, {\lma}, and {\wood}. With some further manipulations, it is possible to show that {\wplcp} will decrease with {\lma} if $\beta_{{\rm kl2}} >1$. Likewise {\wplcp} will decrease with {\wood} if $\beta_{{\rm ks2}} >1$. The parameters $\beta_{{\rm kl2}}$ and $\beta_{{\rm ks2}}$ give the slope relating tissue turnover rates to {\lma} and {\wood}, respectively. Since in this analysis, we have assumed $\beta_{{\rm kl2}} >1$ and $\beta_{{\rm ks2}} >1$, species with low {\lma} and low {\wood} are predicted to be less shade tolerant (Fig. \ref{fig:wplcp}). At low {\lma} ({\wood}), leaf (sapwood) turnover is higher and thus a greater light income is needed to offset increased turnover. {\wplcp} also decreases with height because as size increases, the total amount of carbon needed to offset respiratory and turnover costs in the stem also increases \citep{Givnish-1988}. In addition, {\wplcp} varies with {\nitrogen}. At small sizes, {\wplcp} increases with {\nitrogen} across the band of values typically observed in real plants, i.e. high {\nitrogen} makes seedlings shade intolerant. At larger sizes, as biomass growth declines to zero, {\wplcp} begins to increase again for very low values of {\nitrogen}. All these results  match empirical patterns (Table \ref{tab:phenomena}).

@@

Using a model relating plant physiological function and mass allocation to prominent traits, we have clarified how traits could impact on plant growth across the life cycle. Our approach extends a widely-used theoretical model for seedlings, which links mass-based growth rate to the trait {\lma} \citep{Lambers-1992, Wright-2000}, to explicitly include influences of size, light environment, and other prominent traits. During recent decades, functional traits have captured the attention of ecologists, in large part because of the ability to organise the world's plant species along standard dimensions \citep{Westoby-2002}. However, it has remained unclear how or whether prominent traits influence growth outcomes \citep{Poorter-2008, Wright-2010,Paine-2015}. Matching growing empirical evidence (Table \ref{tab:phenomena}), this study shows when and why the direction or strength of a correlation between traits and growth rate shifts with plant size. Moreover, we show that different traits and trade-offs generate different types of response. Combined with the available empirical evidence, these results demand a shift in our understanding of plant ecological strategies, away from one in which species are thought to have a fixed growth strategy throughout their life \citep[e.g.][]{Grime-1977, Adler-2014, Paine-2015} to one in which traits define a size-dependent growth trajectory \citep{Gibert-2016}. Moreover, we find that growth trajectories and the ranking of traits across them are also moderated by the light environment; while traits that minimise costs of tissue respiration and / or turnover also make plants more shade tolerant (i.e. lower \textsc{wplcp}), as is empirically observed \citep{Messier-1999, Craine-2005, Poorter-2006, Baltzer-2007, Lusk-2008}. The {\plant} model, used here, builds on and extends several related approaches, wherein emergent outcomes such as height, diameter, and mass growth arise from the interaction of different tissues and traits \citep[e.g.][]{Givnish-1988, Makela-1997, Moorcroft-2001}. This approach is quite different to models derived from metabolic scaling theory (\textsc{mst}), which derive everything from a single master ``scaling'' equation for mass growth and have thus far been unable to account for size-dependent changes in the correlation between traits and growth rate \citep{Enquist-1999, Enquist-2007}. Our approach is also fundamentally different from statistically-fitted growth models \citep[e.g.][]{Herault-2011, Ruger-2012, Iida-2014} in that it predicts rather than statistically tests for trait-based effects. Our aim is to both explain observed phenomena (Table \ref{tab:phenomena}) and also generate new hypotheses.

\subsection*{LMA doesn't predict growth rates of large plants}

@@

\subsection*{Generalising to other traits and trade-offs}

@@

\subsection*{Implications for trait-based approaches}
There are some broad implications of our work for our understanding of plant ecological strategies and plant growth.

@@

\subsection*{Plasticity of traits through ontogeny}

@@

\subsection*{Comparison with other frameworks}

@@

A feature distinguishing our approach \citep[see also][]{Falster-2011, Falster-2016, Falster-2017} from most of those mentioned above is the explicit linking to trait-based trade-offs. This link enables demographic outcomes and trade-offs to emerge from underlying phsyiologcal function. While such a linkage was also made by \citep{Moorcroft-2001} in the \textsc{ed} model; analyses using \textsc{ed} have focussed on ecosystem-level outcomes rather than the growth of individual plants. Because of its underlying similarities, we expect the dynamics reported here to be also present within the \textsc{ed} model. Another study \citep{King-1999} connected a model of growth for a single species to the trait {\lma}, and likewise predicted a gradual flattening out of the relationship between {\lma} and growth rate with size (as in Fig. \ref{fig:growth_light_height}). Here, we include an additional cost of increased leaf turnover that further penalised low {\lma} strategies when large.

Perhaps the two most controversial elements of our approach concerns the assumptions about tissue replacement and reproductive allocation. Many vegetation models determine allocation based on net primary production (photosynthesis - respiration), whereas we also subtracted tissues lost via turnover before distributing any surplus. This is because we assume tissues lost via turnover are replaced before mass is allocated to either new growth (i.e. growth that leads to a net increase in $M_{\rm l}, M_{\rm b}, M_{\rm s}$ or $M_{\rm r}$) or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants and perennials, but may not hold for some herbs or annuals, where the switch to reproduction may entail a run-down in the vegetative part of the plant. The second assumption we make is that when mature, plants allocate a substantial fraction of their surplus mass to reproduction. While it remains unclear just how much adult plants might allocate to reproduction, recent reviews suggest the fraction may be high \citep{Thomas-2011, Wenk-2015}. Moreover, a long line of theoretical models indicate that allocation should increase as plants age \citep[reviewed by][]{Wenk-2015}. Currently, reproductive allocation receives little attention in models focussing on production \citep[e.g.][]{Sitch-2008, DeKauwe-2014}. For example in the \textsc{ed} model, a fixed 30\% of net primary production is allocated to reproduction, irrespective of plant size. Yet differences in reproductive allocation offer a clear mechanism explaining correlations between a maximum size and growth rate \citep[e.g.][]{Wright-2010}.

@@

\subsection*{Closing remarks}

@@

\matmethods{
\vspace{-2em}

\subsection*{Parameters}
Parameters were sourced mainly from \citep{Falster-2016} (see SI Appendix, Tables \ref{tab:params_core}--\ref{tab:params_hyper} for values). The only exceptions are: i) parameters affecting the relationships between various size metrics and leaf area, outlined in Table \ref{tab:allometry}a, estimated from Fig \ref{fig:assumptions}, and ii) parameters describing the function for reproductive allocation, where we set $r_{\rm r1}=0.8$ and $r_{\rm r2}=10$, implying a relatively rapid transition to reproduction at {\hmat} (see panel for $\frac{{\rm d}M_{{\rm a}}}{{\rm d}B}$ in Fig. \ref{fig:conceptual}).

\subsection*{Assumptions}
The functional-balance assumptions listed in Table \ref{tab:allometry}a were evaluated using data from the Biomass and Allometry Database \citep[\textsc{baad}][]{Falster-2015b}, which includes records for various size metrics from 21084 individual plants across 656 species. We fit standardised major axis ({\sma}) lines \citep{Warton-2006} to characterise bivariate relationships. We implemented a hierarchical model structure, where the distribution of slopes and intercepts among groups is assumed Gaussian.

\subsection*{Software}
The growth model applied here is implemented as the \texttt{FF16} physiological module within v1.1 of the {\plant} package \citep{Falster-2016} for \textsc{r} \citep{R-2015}. The {\plant} package makes use of supporting packages \texttt{Rcpp} \citep{Eddelbuettel-2013} and the Boost Library for C++\citep{Schaling-2014}, via the package \texttt{BH} \citep{Eddelbuettel-2015}. To encode the trait-based trade-offs described above, we {\plant}'s ability to provide a ``hyper-parameterisation'' function, which causes various parameters to covary with traits (for details see SI).
